---
title: "Data cleaning script"
author: "sbsambado"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(xlsx)
library(dplyr)
library(plyr)
library(vegan)
library(Rcapture)

```

## Dataset 1 -- Tick Density
--> final version of exported datafile =  "Tejon_MeanComparisons_Dataset.xlsx"

Step 1. Merge data
```{r merge data}

threeyrs <- read_csv("~/Google Drive/Tejon_clean/raw_data/Total_Ticks_2016_to_2018.csv")
dim(threeyrs)
oneyr <- read_csv("~/Google Drive/Tejon_clean/raw_data/Tick_19.csv")
dim(oneyr)
# get rid of survery column in threeyr since it wasn't completed in oneyrs 
## Maybe double check survey number doesn't matter
threeyrs <- subset(threeyrs, select = -c(3))

# make sure names are the same
## 2016 - 2018
names(threeyrs) <- tolower(names(threeyrs))

# change column names
names(threeyrs)[6] <- "total"
names(threeyrs)[7] <- "deoc"
names(threeyrs)[8] <- "ipac"
names(threeyrs)[9] <- "deva"
names(threeyrs)[10] <- "other"

## 2019
# make sure names are the same
names(oneyr) <- tolower(names(oneyr))

# change column names
names(oneyr)[6] <- "total"
names(oneyr)[7] <- "deoc"
names(oneyr)[8] <- "ipac"
names(oneyr)[9] <- "deva"
names(oneyr)[10] <- "other"


tick <- rbind(threeyrs, oneyr)

## add plotID

tick <- transform(tick, plotID = as.numeric(interaction(site, plot,block, drop = TRUE)))

```

Step 2. Check data structures
```{r check data structures}
dim(tick) # 1449 11
str(tick)

tick$year <- as.numeric(tick$year)
tick$site <- as.factor(tick$site)
tick$plot <- as.factor(tick$plot)

# change it to Tick cause I had further analyses labeled as Tick vs tick
Tick <- tick

# write to an excel sheet
write.xlsx(Tick, file = "Tejon_MeanComparisons_Dataset.xlsx",
          col.names = TRUE, row.names = FALSE)
```

## Dataset 2 -- Tick, Mammal, Climate Data
--> final version of exported file = "Tejon_MixedModels_Dataset.xlsx"

**OPTIONAL**: data not needed for these analyses but maybe down the road

Step 3. Add additional data to tick counts
~ Lots of movin' and groovin here. Trying to add multiple datasets together while avoiding duplication and too much observation lost ~

a. log of tick
b. climate (temp, precip) 
c. vertebrate (species richness, diversity, pop estimates)
d. celcius
e. temperature difference
f. elevation
g. rain year
h. plotID

```{r add additional data}
## a. log of tick
# log transform data
Tick$log_total <- log(Tick$total + 1)

## b. climate

############## temperature ##############
temp <- read_csv("~/Google Drive/Tejon_clean/raw_data/Tejon_temperature_02022021.csv")
temp.tick <- merge(Tick, temp)
dim(temp.tick) # 1449, 14 (added meanmaxF, meanminF, station)
#View(temp.tick)

############## precipitation ##############

precip <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_tick_climate_10162020.csv")

precip_subset <- subset(precip, select = c(1:5,19,20))
names(precip_subset) <- tolower(names(precip_subset))
temp.tick.precip <- plyr::join(temp.tick, precip_subset)
dim(temp.tick.precip) #2025, 16 (added precip (in), precip (mm))
names(temp.tick.precip)
# seems like data got duplicated let's fix that

# best way is to remove duplicate rows which I lose ~50 observations but I think that is a better way to do it then extract unique elements by columns since alot of these observations are repetitive

temp.tick.prep.short <- temp.tick.precip %>%
  distinct()

dim(temp.tick.prep.short) # 1395, 16

## c. vertebrate

############## mammal metrics ############## 

mam <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mammal_02022021.csv")
names(mam) <- tolower(names(mam))

climate.mam <- join(temp.tick.prep.short, mam)
dim(climate.mam) # 1895, 19 (added cum_mammal, shan_mammal, rich_mammal)

# remove duplicate rows

climate.mam.short <- climate.mam %>%
  distinct()
dim(climate.mam.short) #1395, 19 # same amount of observations after climate dataset


############## pop estimates ############## 

pop_estimates <-read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mamliz_popestimates.csv")
dim(pop_estimates) #27 5

climate.mam.combo <- join(climate.mam.short, pop_estimates)
dim(climate.mam.combo) # 1395   21 (added liz estimates, mam estimates)
names(climate.mam.combo)[20] <- "liz_estimates"

dim(climate.mam.combo) # 1395   21
names(climate.mam.combo)

## d. celcius
climate.mam.combo$mean_maxC <- (climate.mam.combo$mean_maxF - 32)*(5/9)
climate.mam.combo$mean_minC <- (climate.mam.combo$mean_minF - 32)*(5/9)


## e. temperature difference 
climate.mam.combo$tempdifF <- (climate.mam.combo$mean_maxF - climate.mam.combo$mean_minF)
climate.mam.combo$tempdifC <- (climate.mam.combo$mean_maxC - climate.mam.combo$mean_minC)


## f. elevation
  climate.mam.combo$elevation <- climate.mam.combo$station
  climate.mam.combo$elevation[climate.mam.combo$station == "Lebec"] <- 1089.7
  climate.mam.combo$elevation[climate.mam.combo$station == "Loraine"] <- 1290.8
  climate.mam.combo$elevation[climate.mam.combo$station == "Chuchupate"] <- 1493.5

# g. rain year
  # this is a dumb way but it is what it is
  climate.mam.combo$rain_year <- as.numeric(climate.mam.combo$year)
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'April'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'May'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'June'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'July'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'August'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'September'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'October'] <- "2015/2016"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'November'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'December'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'January'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'February'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'March'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'April'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'May'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'June'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'July'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'August'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'September'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'October'] <- "2016/2017"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'November'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'December'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'January'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'February'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'March'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'April'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'May'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'June'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'July'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'August'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'September'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'October'] <- "2017/2018"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'November'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'December'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'January'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'February'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'March'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'April'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'May'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'June'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'July'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'August'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'September'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'October'] <- "2018/2019"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'November'] <- "2019/2020"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'December'] <- "2019/2020"

## h. plotID
climate.mam.combo <- transform(climate.mam.combo, plotidname = as.character(interaction(site, plot,block, drop = TRUE)))
climate.mam.combo <- transform(climate.mam.combo, plotID = as.numeric(interaction(site, plot,block, drop = TRUE)))
str(climate.mam.combo) # 1395 observations, 26 variables (added plotidname, plotID)


```

Step 4. Tidy data (will need to upload optional data from step 3)

```{r tidy data}
# put data in right format
str(climate.mam.combo)

climate.mam.combo$year <- as.factor(climate.mam.combo$year)
climate.mam.combo$block <- as.factor(climate.mam.combo$block)
climate.mam.combo$plotID <- as.factor(climate.mam.combo$plotID)
climate.mam.combo$rain_year <- as.factor(climate.mam.combo$rain_year)
climate.mam.combo$elevation <- as.numeric(climate.mam.combo$elevation)

climate.mam.combo$month <- factor(climate.mam.combo$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
climate.mam.combo$site <- factor(climate.mam.combo$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

climate.mam.combo$plot <- factor(climate.mam.combo$plot, 
                    levels = c('Open', 'Partial', 'Total'))

# rename precip columns
names(climate.mam.combo)[15] <- "precip_in"
names(climate.mam.combo)[16] <- "precip_mm"

```

Step 5. Final data set to use for analyses
--> final version of exported file = "Tejon_MixedModels_Dataset.xlsx"
```{r export file for analyses}
# Finally, the final dataset for GLMMS
FINAL <- climate.mam.combo
str(FINAL) # 1395 observations, 30 variables

# write to an excel sheet
write.xlsx(FINAL, file = "Tejon_MixedModels_Dataset.xlsx",
           col.names = TRUE, row.names = FALSE)

```



### Additional analyses that were used to make input variables for glmms

#### Population estimates for lizards and mammals

Original datasets I was given: `Lizard_Counts_TREE` for lizards and `2018_smammals` and `Small_mammal_captures_19`

From those, I created an excel sheet called `tejon_mamliz_popestimates` that I made based on the analyses in lizard estimates and mammal estimates. Yes, I know this could have been done with a for loop but couldn't get my brain to think about it. This excel sheet is used to build the full data sheet with climate and tick counts

Step 1. Upload & tidy data
```{r data sheets for pop estimates}
# Used these data sheets to calculate pop estimates. All calculations used to make population estimates can be found in the dataset `tejon_mamliz_popestimates`.
### mammals 
mamcounts <- read_csv("~/Google Drive/Tejon_clean/raw_data/mammal_counts_02182021.csv")
dim(mamcounts) # 229   9

#str(mamcounts)
mamcounts$day1 <- as.numeric(mamcounts$day1)
mamcounts$day2 <- as.numeric(mamcounts$day2)
mamcounts$day3 <- as.numeric(mamcounts$day3)
mamcounts$site <- as.factor(mamcounts$site)
mamcounts$plot <- as.factor(mamcounts$plot)
mamcounts$block <- as.numeric(mamcounts$block)
#str(mamcounts)

### lizards
lizcounts <- read_csv("~/Google Drive/Tejon_clean/raw_data/lizardcounts_02172021.csv")
dim(lizcounts) # 1745 , 10
names(lizcounts) <- tolower(names(lizcounts))

#str(lizcounts)
lizcounts$day1 <- as.numeric(lizcounts$day1)
lizcounts$day2 <- as.numeric(lizcounts$day2)
lizcounts$day3 <- as.numeric(lizcounts$day3)
lizcounts$site <- as.factor(lizcounts$site)
lizcounts$plot <- as.factor(lizcounts$plot)
lizcounts$block <- as.numeric(lizcounts$block)
#str(lizcounts)

```


Step 2. Lizard population estimates

```{r lizard estimates}
AridO <- lizcounts %>%
  filter(site == "Arid", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# ARID OPEN BLOCK 1 = 51.35882

AridO <- lizcounts %>%
  filter(site == "Arid", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# ARID OPEN BLOCK 2 = 60.5687

AridO <- lizcounts %>%
  filter(site == "Arid", plot =="Open", block == 3) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# ARID OPEN BLOCK 3 = 82.54301



##########################################################################
AridP <- lizcounts %>%
  filter(site == "Arid", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID PARTIAL Block 1 = 73.4186

AridP <- lizcounts %>%
  filter(site == "Arid", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID PARTIAL Block 2 = 72.96941

AridP <- lizcounts %>%
  filter(site == "Arid", plot =="Partial", block == 3) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID PARTIAL Block 3 = 69.2156


##########################################################################

AridT <- lizcounts %>%
  filter(site == "Arid", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTALBLOCK 1 = 67.28581

AridT <- lizcounts %>%
  filter(site == "Arid", plot =="Total", block == 2) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTALBLOCK 2 = 60.0097

AridT <- lizcounts %>%
  filter(site == "Arid", plot =="Total", block == 3) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTALBLOCK 3 = 88.44651


##############################################################################
##############################################################################
##############################################################################


IntermediateO <- lizcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate OPEN BlOCK 1 = 129.3622

IntermediateO <- lizcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate OPEN BlOCK 2 = 84.6362

IntermediateO <- lizcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 3) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate OPEN BlOCK 3 = 207.9784

##########################################################################
IntermediateP <- lizcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 1) %>%
  select(day1, day2, day3) %>%
  na.omit(lizardcounts)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 1 = 133.5039

IntermediateP <- lizcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 2 = 47.43047

IntermediateP <- lizcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 3) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 3 = 162.613

##########################################################################

IntermediateT <- lizcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance , lowest BIC had negative value with Mbh so I chose next best model Mb abundance = 137.9494

## Intermediate TOTAL BLOCK 1 = 398.0468

IntermediateT <- lizcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 2) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate TOTAL BLOCK 2 = 84.5119

IntermediateT <- lizcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 3) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate TOTAL BLOCK 3 = 176.55


```

Step 3. Mammal population estimates

```{r mammal estimates }
AridO <- mamcounts %>%
  filter(site == "Arid", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID OPEN Block 1 = 2

AridO <- mamcounts %>%
  filter(site == "Arid", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID OPEN Block 2 = 5.236068

AridO <- mamcounts %>%
  filter(site == "Arid", plot =="Open", block == 3) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID OPEN Block 3 = 2.618034

##########################################################################
AridP <- mamcounts %>%
  filter(site == "Arid", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID PARTIAL BLOCK 1= 15.51451
AridP <- mamcounts %>%
  filter(site == "Arid", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance , lowest BIC was negative so chose next lowest model M0

## ARID PARTIAL BLOCK 2= 45.38767
AridP <- mamcounts %>%
  filter(site == "Arid", plot =="Partial", block == 3) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID PARTIAL BLOCK 3= estimates were in the 500000000 so I chose to omit this block 

##########################################################################

AridT <- mamcounts %>%
  filter(site == "Arid", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTAL block 1= 15.51451

AridT <- mamcounts %>%
  filter(site == "Arid", plot =="Total", block == 2) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTAL block 2= 40.80791

AridT <- mamcounts %>%
  filter(site == "Arid", plot =="Total", block == 3) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## ARID TOTAL block 3= 15.51451


##############################################################################
##############################################################################
##############################################################################

# no mammals trapped at block 1 intermediate open

IntermediateO <- mamcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate OPEN BLOCK 2 = 5.236068

IntermediateO <- mamcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 3) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate OPEN BLOCK 3 = estimates were in the 7e09 so I chose to omit this block 


##########################################################################
IntermediateP <- mamcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 1= 7.734787

IntermediateP <- mamcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 2 = 2

IntermediateP <- mamcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 3) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Intermediate PARTIAL BLOCK 1= estimates were in 700 so omitted this data

##########################################################################

IntermediateT <- mamcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] 

## Intermediate TOTAL BLOCK 1= 	31.33333

IntermediateT <- mamcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 2) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] 

## Intermediate TOTAL BLOCK 2= 	estimates over 5 mil so omitted

IntermediateT <- mamcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 3) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] 
## Intermediate TOTAL BLOCK 1= 	2


#############################################################################
# MESIC
#############################################################################
#############################################################################

MesicO <- mamcounts %>%
  filter(site == "Mesic", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Mesic OPEN BLOCK 1= 28.5

MesicO <- mamcounts %>%
  filter(site == "Mesic", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Mesic OPEN BLOCK 2= 2

MesicO <- mamcounts %>%
  filter(site == "Mesic", plot =="Open", block == 3) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Mesic OPEN BLOCK3 = 12.0847


##########################################################################
MesicP <- mamcounts %>%
  filter(site == "Mesic", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(MesicP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# MESIC PARTIAL BLOCK1 = 215.727

MesicP <- mamcounts %>%
  filter(site == "Mesic", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(MesicP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# MESIC PARTIAL BLOCK 2= 15.51451

MesicP <- mamcounts %>%
  filter(site == "Mesic", plot =="Partial", block == 3) %>%
  select(day1, day2, day3)
  
abundanceP = closedp.t(MesicP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceP$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

# MESIC PARTIAL BLOCK = estimates were 10mil so omitted them

##########################################################################

MesicT <- mamcounts %>%
  filter(site == "Mesic", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(MesicT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance # second best fit model is M0, since 1st best fit is -10 with Mb

## Mesic TOTAL BLOCK 1 = 62.26667

MesicT <- mamcounts %>%
  filter(site == "Mesic", plot =="Total", block == 2) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(MesicT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance # second best fit model is M0, since 1st best fit is -10 with Mb

## Mesic TOTAL BLOCK 2 = 7.854102

MesicT <- mamcounts %>%
  filter(site == "Mesic", plot =="Total", block == 3) %>%
  select(day1, day2, day3)
  
abundanceT = closedp.t(MesicT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceT$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance # second best fit model is M0, since 1st best fit is -10 with Mb

## Mesic TOTAL BLOCK 1 = 17.20092

```
#### Mammal metrics of Shannon richness and species diversity

*Original datasets I was given: `2018_smammals` and `Small_mammal_captures_19`* 

From those, I created excel sheets called `tejon_mammals1819_09022020` and `tejon_mammalcount_09032020`.


```{r mammal metrics, warning=FALSE}
# full dataset
mammals <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mammals1819_09022020.csv")
# subset for important
mammal <-data.frame(mammals[,4:6],mammals[,9])


# numeric dataframe for community analysis
Count <- read_csv("~/Google Drive/Tejon_clean/raw_data/tejon_mammalcount_09032020.csv")

# remove the arid with NAs, this will be the full dataset with species richenss and diversity added
count <- Count[-4,]
# keep only numeric data to run community analyses
numeric <- count[,c(-1,-2,-9,-10)]
# organize data into right forms
count$plot <- factor(count$plot, 
                        levels = c('open','partial','total'))
# species richness
sppr <- specnumber(numeric)
# shannon diversity
shannondiv <- diversity(numeric)
# add species richness to count
count$sppr<-specnumber(numeric)
# add shannon diversity to count
count$shannondiv <- diversity(numeric)

aggregate(shannondiv ~plot, data = count, FUN = mean)
```

