---
title: "BestAnalyses_script"
author: "sbsambado"
date: "10/26/2021"
output: html_document
---

Set up
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(car)
library(knitr)
library(lme4)
library(jtools)
library(effects)
library(psych)
library(gt)
library(MuMIn)


```

*Notes*
plot = cattle
site = climate

## Question 1: How do environmental stressors - cattle & climate - impact tick counts? 

--> best fit model incorporates both stressors
--> however, looked at individually, climate is the more strong predictor of tick counts
--> best fit random effects were month & plotID  

Step 1. Import data set
```{r import}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")
```

Step 2. Build best fit models
- look at Tejon_Analyses_20211021 for all model combinations

```{r best fit models}
mod_plot <- glmer.nb(total ~ plot + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")

mod_site <- glmer.nb(total ~ site  + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")

mod_both <- glmer.nb(total ~ site + plot + (1|month) + (1|plotID), 
            data = Tick, family = "poisson")


model_results <- AIC(mod_plot, mod_site, mod_both)

model_models <- list(mod_plot, mod_site, mod_both)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)
for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$variables <- c("plot", "site", "plot + site")

model_models <- kable(model_results, digits = 2, align = "c") %>%
  kable_styling()

model_models

```

Step 3. Model diagnostics

3A. summary functions
```{r model diagnostics pt 1}

## summary()
summary(mod_plot) # p = 0.00770
summary(mod_site) # p = 0.00567
summary(mod_both) # p = 0.000187

## summ() function
               # Est    S.E.    z value   p
summ(mod_plot) # -2.05   0.77    -2.67   0.01  ; Pseudo-R² (total) = 0.81 
summ(mod_site) # -2.09   0.75    -2.77   0.01  ; Pseudo-R² (total) = 0.81 
summ(mod_both) # -2.82   0.76    -3.74   0.00  ; Pseudo-R² (total) = 0.81

```

3B. Akaike weight
- get relative support for each model by standardizing the likelihood of the model
```{r model diagnostics pt 2}

## Option 1: Calculate weights manually

# calculate deltai2

delta1 <- exp(-0.5*(AIC(mod_plot) - AIC(mod_both)))
delta2 <- exp(-0.5*(AIC(mod_site) - AIC(mod_both)))
delta3 <- exp(-0.5*(AIC(mod_both) - AIC(mod_both)))

# sum likelihoods
sum_likes <- sum(c(delta1, delta2, delta3))

# calculate weights
weight1 <- delta1/sum_likes
weight2 <- delta2/sum_likes
weight3 <- delta3/sum_likes

# lets make a table
weights <- tribble(
  ~Model,   ~Delta,           ~Weights,   ~Variable,
  "mod_plot", round(delta1, 2), round(weight1, 2), "plot",
  "mod_site", round(delta2, 2), round(weight2, 2), "site",
  "mod_both", round(delta3, 2), round(weight3, 2), "both") %>% 
  gt() %>% 
  tab_style(
    style = cell_text(size = "small"),
    locations = cells_body())


weights


## Option 2: dredge the model

min(Tick$total)
model <- na.omit(Tick)
big.model <- glm(total  ~ plot + site + plot*site, 
                 family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  tab_header(title="Model Selection",
             subtitle = "total ~ plot + site + plot*site")

glm.best

# delta = 0 is best model

```

Step 4. Plot model outputs
```{r plot model outputs, results = 'asis'}

## plot_summs()
plot_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           model.names = c("Only Cattle (plot)", "Only Climate (site)", "Both Cattle & Climate"),
           inner_ci_level = .8,
           robust = TRUE,
           coefs = c("Partial plot" = "plotPartial",
                     "Total plot" = "plotTotal",
                     "Inter. site" = "siteIntermediate", 
                     "Mesic site" = "siteMesic"),
           legend.direction = "horizontal")

```

Step 5. Create output table of model results
```{r model table statistics}
export_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           error_format = "({statistic}, 95% CI [{conf.low}, {conf.high}], p = {p.value})",
             to.file = "docx", 
             file.name = "Tejon_BestFitModels.docx")
```


## Question 2: How do environmental stressors impact populations of two important vertebrate hosts -- lizards and mammals -- in the tick system?

--> lizards are significantly more impacted by climate than plot
--> mammals are trending to be significantly impacted by plot than climate

Step 1. Import & tidy dataset
```{r import and tidy}

pop <- read_csv("raw_data/tejon_mamliz_popestimates.csv")

# rename columns to be consistant
names(pop)[5] <- "mam_estimate"

# change data structure for analyses
pop$site <- as.factor(pop$site)
pop$plot <- as.factor(pop$plot)

```

Step 2. Compare means of population estimates

2A. Lizards
```{r lizards}
# lizards were only sampled at two sites, so will use a two sample t-test if normal

# check normality
qqPlot(pop$liz_estimate) # looks normal, but..
shapiro.test(pop$liz_estimate) # p < 0.05, so data are not normal

# check variance
leveneTest(pop$liz_estimate ~ pop$site) #  p < 0.05, F = 7.4; variances are different

# non-parametric Mann-Whitney U is testing the shape of the two distributions are different

# wilcox rank sum test with continuity correction (non-equal variances)
wilcox.test(liz_estimate ~ site, data = pop) #W = 11, p-value  = 0.01306

# p < 0.05, reject H0 that means are equal to 0

## lizard population estimates are different at Intermediate and Arid sites

```

2B. Mammals
```{r mammals}

# mammal populations did not significantly differ across sites or plots but that may be due to low sample numbers with kruskal wallis

# let's do open vs total plots --> mann-whitney U test

pop_sub <- subset(pop, pop$plot != "Partial")
wilcox.test(mam_estimate ~ plot, data = pop_sub)
# W = 11, p-value = 0.05488 (trending significant)
```

## Question 3: Are tick counts impacted by lizard and mammal population estimates?

--> lizards population estimates are more associated with tick counts than mammals

Step 1. Import & tidy dataset
```{r import and tidy}
FINAL <- read_csv("data_analyses/Tejon_MixedModels_Dataset.csv")

# drop N/A values
FINAL <- FINAL[which(FINAL$mam_estimates != "#N/A"),]
FINAL <- FINAL[which(FINAL$liz_estimates != "#N/A"),]

# check it worked
unique(FINAL$mam_estimates)
unique(FINAL$liz_estimates)

str(FINAL)
```

Step 2. Spearman's rank correlation

```{r spearman rank correlation}
# select subset of data to use for test & visualize
FINAL_cor <- subset(FINAL, select = c(6,20:21))

# change structure of data for analyses
FINAL_cor$total <- as.numeric(FINAL_cor$total)
FINAL_cor$mam_estimates <- as.numeric(FINAL_cor$mam_estimates)
FINAL_cor$liz_estimates <- as.numeric(FINAL_cor$liz_estimates)

# visualize data
pairs.panels(FINAL_cor)

## run correlation test

# lizards
cor.test(FINAL_cor$total,FINAL_cor$liz_estimates,
         method = "spearman")
# p-value = 1.755e-05
# rho = 0.1687151

# mammals
cor.test(FINAL_cor$total,FINAL_cor$mam_estimates,
         method = "spearman")
# p-value = 0.3669
# rho = -0.03569831 


```

Step 3. Build models : GLM with Poisson distribution 

~ I think this may not be appropriate based on the aggregate of the pop estimates ~

```{r glm with poisson}

mod_mam <- glm(total ~ mam_estimates, data = FINAL,
               family = "poisson")

mod_liz <- glm(total ~ liz_estimates, data = FINAL,
               family = "poisson")

mod_ml <- glm(total ~ mam_estimates + liz_estimates, data = FINAL,
               family = "poisson")

## Model comparison

model_results <- AIC(mod_mam, mod_liz, mod_ml)

model_models <- list(mod_mam, mod_liz, mod_ml)

model_results$BIC <- sapply(model_models, BIC)
model_models_summary <- lapply(model_models, summary)

for(i in 1:(length(model_models)-1)){ #this creates a variable i that starts with the value i=1
  model_results$rsq[i] <- model_models_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  model_results$adj_rsq[i] <- model_models_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

model_results$variables <- c("mam", "liz", "mam + liz")

model_models <- kable(model_results, digits = 2, align = "c") %>%
  kable_styling()

model_models
```

Step 4. Model diagnostics

4A. summary functions
```{r model diagnostics pt 1}

## summary()
summary(mod_mam) # p = <2e-16
summary(mod_liz) # p = <2e-16
summary(mod_ml) # p = <2e-16

## summ() function
               # Est    S.E.    z value   p
summ(mod_mam) #  0.91   0.04    23.12   0.00   
summ(mod_liz) # -1.46   0.09   -17.04   0.00 
summ(mod_ml) #  -1.26   0.10   -12.49   0.00

```

4B. Dredge model
```{r model diagnostics pt 2}
## Option 2: dredge the model

min(FINAL$total)
model <- na.omit(FINAL)
big.model <- glm(total ~ mam_estimates + liz_estimates, 
               family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  tab_header(title="Model Selection",
             subtitle = "total ~ mam + liz")

glm.best

```

