---
title: "Data Figures and Tables"
author: "sbsambado"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(ggfortify)
library(gtsummary)
library(lme4)
library(effects)
library(jtools)
library(huxtable)

# set up some visualization vectors
clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('brown1','tan1','lightgoldenrod1')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))

ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))

xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))


```

## Figure 1
**Tick counts by one variable (plot or site)**

```{r}
Tick <- read_csv("Viz_Tejon_MeanComparisons_Dataset.csv")

# omit 0s to make plots look nicer, but this may not be kosher
Tick_no0 <- Tick[which(Tick$total != 0),]

plot1 <- ggplot(Tick_no0, aes(x = plot, y = log(total +1), fill = plot))+
  #geom_jitter(alpha = .15, width = .2)+
  geom_boxplot()+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  guides(fill = FALSE)+
  annotate('text', label='A', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

site1 <-ggplot(Tick_no0, aes(x = site, y = log(total +1), fill = site))+
  #geom_jitter(alpha = .15, width = .2)+
  geom_boxplot() +
  theme_bw() +
  labs(x = 'Site', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)+
  annotate('text', label='B', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

var_plot <-ggarrange(plot1 + rremove('ylab'), site1 + rremove('ylab'),
          ncol = 2)
annotate_figure(var_plot,
          left = text_grob("Log of tick counts (2016 - 2019)", rot = 90, face = "bold"))
#ggsave("data_visualizations/Fig1_OneVarTick.jpeg", dpi = 320)          


```
## Figure 2. 
** Tick counts by two variables (plot and site)**

```{r}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

# omit 0s to make plots look nicer, but this may not be kosher
Tick_no0 <- Tick[which(Tick$total != 0),]

ggplot(Tick_no0, aes(x = site, y = log(total + 1), fill = plot))+
  geom_boxplot()+
  #geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Site', y = "Log of tick counts (2016 - 2019)", fill = 'Plot') +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = c(.911, .816), # horizontal, verticle
        legend.box.background = element_rect(color = 'black', fill =NA),legend.key.size = unit(.8,"line"))
ggsave("data_visualizations/Fig2_TwoVarTick.jpeg", dpi = 320)  

```

## Figure 3. 
**Climate by site**

```{r}
FINAL <- read_csv("data_analyses/Tejon_MixedModels_Dataset.csv")

# change data structure
FINAL$site <- as.factor(FINAL$site)
FINAL$plot <- as.factor(FINAL$plot)
FINAL$tempdifC <- as.numeric(FINAL$tempdifC)
FINAL$mean_maxC <- as.numeric(FINAL$mean_maxC)

ggplot(FINAL, aes(y = tempdifC, x = site, fill = site))+
  geom_boxplot() +
  #geom_jitter(alpha = .15, width = .2)+
  theme_bw()+
  labs(x = 'Site', y = 'Difference in monthly\n mean temperature (°C)', fill = "Site")+
  xtitletheme + ytitletheme +
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)
ggsave("data_visualizations/Fig3_DiffTempCSite.jpeg", dpi = 320)

```
## Figure 4. 
**Effects plot: tick counts ~ cattle + climate**

```{r}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

# mod_plot <- glmer.nb(total ~ plot + (1|month) + (1|plotID), 
#             data = Tick, family = "poisson")
# 
# mod_site <- glmer.nb(total ~ site  + (1|month) + (1|plotID), 
#             data = Tick, family = "poisson")
# 
# mod_both <- glmer.nb(total ~ site + plot + (1|month) + (1|plotID), 
#             data = Tick, family = "poisson")

## plot_summs()
plot_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           model.names = c("Only Cattle (plot)", "Only Climate (site)", "Both Cattle & Climate"),
           inner_ci_level = .8,
           robust = TRUE,
           coefs = c("Partial plot" = "plotPartial",
                     "Total plot" = "plotTotal",
                     "Inter. site" = "siteIntermediate", 
                     "Mesic site" = "siteMesic"))
ggsave("data_visualizations/Fig4_Coefficients.jpeg", dpi = 320)


```
### Supplemental

## Figure 5.
**Tick counts by month (site)**

```{r}

Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Tick$site <- factor(Tick$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

Tick$plot <- factor(Tick$plot, 
                    levels = c('Open', 'Partial', 'Total'))

ggplot(Tick, aes(x = site, y = log(total +1), fill = site))+
  facet_wrap(~ month, ncol = 6)+
  geom_boxplot()+
  theme_bw() +
  scale_x_discrete(labels = c("A", "I", "M")) +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'))+
  labs(x = 'Site', y = 'Log of tick counts (2016 - 2019)') +
  xtitletheme + ytitletheme +
  guides(fill = FALSE)+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))
ggsave("data_visualizations/Fig5_MonthlyTickSite.jpeg", dpi = 320)



```

## Figure 6. 
**Tick counts by month (both variables)**
```{r}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Tick$site <- factor(Tick$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

Tick$plot <- factor(Tick$plot, 
                    levels = c('Open', 'Partial', 'Total'))

ggplot(Tick, aes(x = site, y = log(total +1), fill = plot))+
  facet_wrap(~ month, ncol = 6)+
  geom_boxplot()+
  theme_bw() +
  scale_x_discrete(labels = c("A", "I", "M")) +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'),
        legend.position =  'bottom')+
  labs(x = 'Site', y = 'Log of tick counts (2016 - 2019)') +
  xtitletheme + ytitletheme +
  #guides(fill = FALSE)+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))
ggsave("data_visualizations/Fig6_MonthlyTickSitePlot.jpeg", dpi = 320)
```

## Figure 7.
**Vertebrate population estimates**

```{r}
pop <- read_csv("raw_data/tejon_mamliz_popestimates.csv")

# rename columns to be consistant
names(pop)[5] <- "mam_estimate"

# change data structure for analyses
pop$site <- as.factor(pop$site)
pop$plot <- as.factor(pop$plot)

liz <- ggplot(pop, aes(x = site, y = liz_estimate, fill = site))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  labs(x = "Site", y = "Lizard Pop. Estimates")+
  guides(fill = FALSE)

# omit outlier
pop_sub <- pop[which(pop$mam_estimate != "216"),]

mam <- ggplot(pop_sub, aes(x = plot, y = mam_estimate, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  labs(x = "Plot", y = "Mammal Pop. Estimates")+
  guides(fill = FALSE)

ggarrange(liz, mam)
ggsave("data_visualizations/Fig7_VertPopEstimates.jpeg", dpi = 320)

```
## Figure 8. 
**Dredge of best fit tick count models**
```{r}

Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

min(Tick$total)
model <- na.omit(Tick)
big.model <- glm(total  ~ plot + site + plot*site, 
                 family = "poisson",
                 data = model,
                 na.action = na.pass)
sal.dredge <- dredge(big.model)

glm.best <-model.sel(sal.dredge) %>% 
  gt() %>% 
  fmt_number(columns = 6:9,
             decimals = 2) %>%
  tab_header(title="Model Selection",
             subtitle = "tick count ~ plot + site + plot*site") 

glm.best %>%
  gtsave("data_visualizations/Fig8_ModelSelection.png")
```

## Figure 9
**Model outputs for GLMM negative binomial**
```{r}
export_summs(mod_plot, mod_site, mod_both,
           scale = TRUE,
           error_format = "{statistic}, 95% CI [{conf.low}, {conf.high}]",
           borders = .5,
  outer_borders = 0,
             to.file = "pdf", 
             file.name = "data_visualizations/Fig9_BestFitModelsQ1.pdf")

```

## Figure 10.
**summaries of tick counts by site and plot**
```{r}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")

counts <- Tick %>%
  group_by(site, plot) %>%
  tally(total) 


pdf("data_visualizations/Fig10_TickCount.pdf")
grid.table(counts)

```

## Figure 11.
**summaries of tick counts by month**
```{r}
Tick <- read_csv("data_analyses/Tejon_MeanComparisons_Dataset.csv")
Tick$month <- factor(Tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))

counts_month <- Tick %>%
  group_by(month) %>%
  tally(total) 


pdf("data_visualizations/Fig11_TickCountMonth.pdf")
grid.table(counts_month)
```

